### 2020-12-10 msc-thesis.tex
New Introduction

\chapter{Context}

Since invention of the World Wide Web, the number of its users increased rapidly. Today, more than 90 percent of the German population use the Internet \cite{Onlinestudie}. Enterprises recognised this potential to inform and interact with a large number of people. Today, an often used method can be called "digital customer self-service" (DCSS). Specialized software tools enable customers to use services of an enterprise through the internet without direct human interaction.

Developing and maintaining these tools and the underlying architecture can be a difficult and expensive task. Required software solutions can therefore be accessed from commercial DCSS providers. In order to use their tools, enterprises will have to integrate them into existing system architectures. Depending on for example the complexity of the tools, different integration approaches can be applied.

@@

An evolutionary step in customer profile management is the usage of self-sovereign identities (SSI): instead of each enterprise, the customer provides a user profile. This profile can be managed by a trusted DCSS provider for SSI. In order for this concept to work, each enterprise will have to integrate the provider into their existing user profile systems. The result would be one user profile for one person which identifies him as a customer for multiple integrated enterprises.

A currently relevant example for usage of DCSS with particular interesting requirements for user profiles is the "Online Access Law" (OZG). It requires all administrative services of the German federal republic, each member state and commune to be digitally available through interoperable user profiles. The current plan is to make the profile only available for usage in context of the OZG. From a user perspective, this would be yet another profile to manage.

\chapter{Objective}

Based on the prominent example of the OZG, the bachelor thesis will construct a message based integration architecture which enables system architectures with existing user profiles to accept SSI through its providers.

In the beginning, the terms DCSS, DCSS provider, SSI, SSI provider, SSI connector and OZG are defined and the underlying concept is explained.

Then, preparatory work relevant for the integration architecture is presented. Based on process diagrams of digitalized administration services, several use cases of DCSS are listed and evaluated for their relevance. They provide important information on which functionalities the SSI provider is expected to have and the integration architecture is required to integrate.

As an integration approach through a connector is used, concrete functionalities and interfaces of a possible connector are specified for usage by the integration architecture.

@@

The integration architecture integrates previously documented connector and components of the system architecture in order to make the selected use cases available through the SSI provider. Integration patterns are used to formulate the integration architecture. In order to save investments the integration reuses as many systems as possible and modifies as few systems as necessary.

The integration architecture is presented iteratively. Each iteration contains increasingly demanding requirements along with the documentation of the increasingly complex integration architecture. Each documentation is done through textual explanations, multiple flow diagrams and one messaging architecture.

@@

\chapter{Fundamentals}

\section{Digital Customer Self-Service}
This chapter introduces the concept of digital customer self-service. It defines DCSS, describes what it is by usage of examples and mentions benefits such as challenges.

\textbf{Definition}: Digital self-service is the consumption of services provided by an enterprise over the internet without human to human interaction.
It is important to distinguish between "service" and "self-service".
The difference is shown through two examples:

\textbf{Example 1}: A service is the availability to purchase items whereas digital self-service is the placement of orders over a website.

\textbf{Example 2}: A service is the availability of video on demand whereas digital self-service is the access of video files over a website.

Digital self-service is one of many possible ways to access a service: An order placement could also be done through email or phone call and access to videos could be granted through same-day-delivery of DVDs.

Customers are able to do digital self-service because they are provided with digital self-service tools. Creating those is not trivial and is the main business model of many companies. Video on demand websites for example do not only flourish as a result of providing access to videos (this business model existed long before them) but as a result of the digital self-service tools they provide.

There exist a lot of different use cases of DCSS. Many can be differentiated into two categories: DCSS for information and DCSS for transaction.
The category DCSS for information contains use cases where DCSS enables users to \textbf{retrieve} information relevant for them. The category DCSS for transaction contains use cases where DCSS enables users to permanently \textbf{add} information to a system, usually in order to trigger some process.

A use case for retrieving information through DCSS are so called "WiKi" pages. Similar to the famous website "Wikipedia", information about for example the usage of a product can be made accessible to customers. In this case, it is not necessary any more to call a support hot-line or visit a shop. Another example are online learning platforms which provide customers with necessary information to for example learn a language. In this case, it is not necessary any more to meet teachers. Online job hunting websites enable its users to find information about job offers which fit their requirements. In this case it is not necessary any more to go to employment bureaus.

A common use case for transaction through DCSS are online retailers which enable users to start payment and delivery processes by providing the necessary information. They are often accompanied by DCSS for information in order to find and select products. In this case it is not necessary any more to go to the local shopping mall. Another use case is the management of user identities where DCSS enables users to authenticate themselves and manage personal information. In this case it is not necessary any more to show up in person at for example a store in order to identify ones identity and give personal information about e.g. the living address.

Reasons for the widespread usage of digital self-service are significant benefits in ease of service usability for customers and cost reduction for enterprises. Customers can self-service from almost every place, easily manage customer data like phone number and address, configure services like subscription plans and access lots of information on available products or trouble shooting steps in an immediate way. Digital customer self-service improves usability, saves time, increases availability, saves money and therefore increases customer experience.
Enterprises, especially in progressive countries, aim to keep employment at a minimum, as it is a big cost factor. Digital self-service turns customers into unpaid employees.

Decreasing employment cost while increasing customer experience sounds like a win-win situation. Sadly, digital self-service does have its downsides. One downside is most immediate to the elderly. In order to use DCSS, computers have to be available and one has to be experienced in their usage. Technological complexity, however, is not only a problem for customers but also for enterprises providing DCSS. Based on existing system architectures, adding DCSS functionalities can turn out to be a difficult task. Some form of integration is often necessary. With increased system complexity and business dependability come additional maintenance cost.

As today most large enterprises provide DCSS, it can be assumed, that at a given size of an enterprise (measured in number of customers), the benefits of DCSS surpass its disadvantages.

Governments see digitalization as a chance for improvements in many important social and economical areas. The process of digitalization can in some cases be described as reinventing existing solutions for usability with computers. One of the goals is to make solutions more easily accessible for users and as a method, provide digital self-service. Governmental regulations relating to digitalization therefore often encourage or demand enterprises and organizations to provide DCSS.

\subsection{DCSS Provider}
As described in the previous section, digital customer self-service is a widely used method on the free market to increase customer experience. In order for customers to use self-service, companies and institutions have to provide necessary tools. Depending on the business case, different categories of DCSS are relevant. Depending on the use case, DCSS is enabled through a variety of software tools.

A chain of stores, for example, might provide DCSS tools for information which enable users to inform themselves about available products and opening hours. One tool could for example provide a search functionality for stores in the local area. Another tool could provide a catalogue of available products when a store is selected. The catalogue could also enable users to filter for relevant products.

Online retailers however will additionally require DCSS tools for transaction. This can be for example be tools for management of user profiles in order for users to authenticate and manage personal information like payment and delivery details.

As just shown, depending on the use case, complexity of technology providing DCSS can be very high. Besides enabling DCSS functionality, technology in use also needs to be maintained in order to adhere to most recent security and data protection standards. Those can also differ between countries. Therefore, in an effort to save costs or to improve functionality companies and organizations could decide to use DCSS software solutions from companies which have specialised in this area: DCSS providers.

As there exist many use cases for DCSS, different software solutions from DCSS providers are possible. The bachelor thesis focuses on DCSS providers for management of identities.

Identity management is important for many businesses and organizations. In order to provide services to individual persons it is necessary to verify their identity, access and update related information and be able to interact. Interaction and verification of identities is straight forward if somebody shows up in person. Personal information of a customer can be managed trough paper files.

DCSS providers for identity management can improve this process for customers and businesses. The following is a possible solution for identity management through DCSS providers, the bachelor thesis will assume.

Description of IDAS ...

- Areas of DCSS: identity, profile, relations, forms, messages, inbox, applications,

- relations, relationship template, app, server, ....

@@

\section{SSI Connector}
seperat wegen bedeutung für Arbeit

@@

\section{OZG DCSS Scenarios}

The information platform of the OZG contains information about many administration services in form of process diagrams \cite{BMI:Ergebnisse}. They serve as a resource to identify important DCSS scenarios of administration processes. The following scenarios were identified:

@@

\section{DCSS Connector}

\section{OZG System Architecture}
This section describes the system architecture of a member state relevant for the execution of the OZG an integration of a DCSS provider. Due to a lack of detailed information, the system architecture is described by components responsible for a category of tasks. Each component is associated to data objects it processes. The definition of components and data objects are based on research about planned system architectures of member states.

@@

This chapter presents a message based integration architecture which enables access to DCSS features of member states through a DCSS provider. Multiple versions of an integration architecture are presented which fulfill increasingly complex requirements. Requirements are formulated as user scenarios which should be accomplishable through the DCSS provider.
Each version is presented in the form of multiple flow charts, each describing sequential steps while carrying out a required scenario and a corresponding messaging infrastructure.

@@

\subsection{Requirements}
In the first version of the integration architecture, users of the DCSS provider should be able to
\begin{itemize}
    \item fill in an application
    \item authenticate
    \item receive and send messages
\end{itemize}

@@

\begin{itemize}
     \item Expanding integration architecture with capabilities for DCSS scenarios of different areas than online administration self-service
     \begin{itemize}
          \item online payment self-service (PayPal, psd2)
          \item online shopping self-service (Amazon)
          \item online health care self-service (DVG)
          \item online entertainment self-service (Netflix, YouTube, Spotify)
          \item online information / support self-service (StackOverflow)
     \end{itemize}
\end{itemize}

>>>>

\chapter{Context}
Since invention of the World Wide Web, the number of its users increased rapidly. Today, more than 90 percent of the German population use the Internet \cite{Onlinestudie}. Enterprises and organizations recognised this potential to inform and interact with a large number of people. A frequently used method is "online self-service" (OSS). Specialized software tools enable customers to use services through the internet without direct human interaction.

Developing and maintaining these tools and the underlying architecture can be a difficult and expensive task. Required software solutions can therefore be accessed from commercial OSS providers. In order to use their tools, enterprises will have to integrate them into existing system architectures. Depending on for example the complexity of the tools, different integration approaches can be applied.

@@

An evolutionary step in customer profile management is the usage of self-sovereign identities (SSI): instead of each enterprise, the customer provides a user profile. This profile can be managed by a trusted OSS provider for SSI.

In order for this concept to work, each enterprise will have to integrate the provider into their existing user profile systems. The result would be one user profile for one person which identifies him as a customer for multiple integrated enterprises.

A currently relevant example for usage of OSS with particular interesting requirements for user profiles is the "Online Access Law" (OZG). It requires all administrative services of the German federal republic, each member state and commune to be digitally available through interoperable user profiles. The current plan is to make the profile only available for usage in context of the OZG. From a user perspective, this would be yet another profile to manage.

\chapter{Objective}
Based on the prominent example of the OZG, the bachelor thesis will construct a message based integration architecture which enables system architectures with existing user profiles to accept SSI through its providers.
The goal of the integration architecture in the example is to make important OZG scenarios accessible with usage of self-sovereign identities instead of the existing user profiles.
The two points of integration are on one hand-side the system architecture relevant for the OZG in form of a list of system components and on the other hand-side the SSI provider in form of a connector.
Integration patterns are used to formulate the integration architecture which enables to use a combination of established and tested messaging solutions. In order to save investments the integration reuses as many system components as possible and modifies as few system components as necessary.

\chapter{Structure of Work}

In the beginning, the terms OSS, OSS provider, SSI, SSI provider, SSI connector and OZG are defined and the underlying concept is explained.

Then, preparatory work relevant for the integration architecture is presented. Based on process diagrams of digitalized administration services, several use cases of OSS are listed and evaluated for their relevance. They provide important information on which functionalities the SSI provider is expected to have and the integration architecture is required to integrate.

An integration approach through a connector is used. Concrete functionalities and interfaces of a possible connector are specified for usage by the integration architecture.

@@

The integration architecture integrates previously documented connector and components of the system architecture in order to make the selected use cases available through the SSI provider. The architecture is presented iteratively. Each iteration contains increasingly demanding requirements along with the documentation of the increasingly complex integration architecture. Each documentation is done through textual explanations, multiple flow diagrams and one messaging architecture.

@@

\chapter{Fundamentals}
The goal of this chapter is to define terminology and to explain basic concepts for better understanding of context and objective of the bachelor thesis.

\section{Online Self-Service}
Self-service has existed for a long time. It generally means the consumption of some service without human interaction but rather in a self-reliant way. One example which is still relevant today are restaurants offering buffets. Instead of a waiter, customers take food for themselves - in self-service. In spring, some farmers allow people to pick flowers from the acre in exchange for a small fee.

In recent decades technology made advanced methods of digital self-service possible. Money can be withdrawn from local ATMs and train tickets can be bought from ticket machines. Technological advancements, at least directed to the average consumer, can be seen as advancements in self-service capabilities. Consumers buy technology which makes them independent from the help of other people.

Since the internet enabled the connection of a very large amount of people, especially the world wide web made it accessible to the general public. The world wide web enables any user to access and interact with technological systems all over the world through a user friendly visual interface. Today, many of those systems are provided by enterprises and other organizations for customers and users to access their services - they enable customers online self-service (OSS).

Online self-service is defined as: self-reliant consumption of services over the internet without human to human interaction.

Depending on the reasoning most websites on the WWW can be described as providing online self-service. OSS therefore rather describes a perspective on the WWW than a use case.

Tow categories of OSS can be differentiated: The information category contains cases where users are enabled to retrieve relevant information. The transaction category contains cases where users modify the systems they interacts with to for example save data or trigger some internal business process.

An example for retrieving information through OSS are so called "WiKi" pages. Similar to the famous website "Wikipedia", information about for example the usage of a product can be made accessible to customers. In this case, it is not necessary any more to call a support hot-line.
Another example are online learning platforms which provide customers with necessary information to for example learn a language. In this case, it is not necessary any more to meet teachers.
Online job hunting websites enable its users to find information about job offers which fit their requirements. In this case it is not necessary any more to visit employment bureaus.

A common use case for transaction through OSS are online retailers which enable users to start payment and delivery processes. They are often accompanied by OSS for information in order to find and select products. In this case it is not necessary any more to go to the local shopping mall.
Another use case is the management of user identities which enables users to authenticate themselves and manage personal information. In this case it is not necessary any more to show up in person at for example a store in order to identify ones identity and give personal information about for example the living address.

Reasons for the widespread usage of online self-service are significant benefits in ease of service usability for customers and cost reduction for enterprises. Customers can self-service from almost every place, easily manage customer data like phone number and address, configure services like subscription plans and access lots of information on available products or trouble shooting steps in an immediate way. Online customer self-service improves usability, saves time, increases availability, saves money and therefore increases customer experience. Enterprises, especially in progressive countries, aim to keep employment at a minimum, as it is a big cost factor.

Digital self-service turns customers into unpaid employees.

\subsection{OSS Provider}
In order for customers to use online self-service, companies and institutions have to provide the necessary software tools. Depending on category and use case of OSS, different software tools exist.

A chain of stores might provide tools which enables the users to inform themselves about available products and opening hours. This could for example be a search functionality for stores in the local area and a catalogue of available products of a selected store. Online retailers however will require additional tools which for example enable management of user profiles to authenticate and communicate with customers.

OSS providers are defined as: enterprises which develop and maintain solutions for OSS and make them accessible to system architectures of customers.

Depending on the complexity of the solution, different methods of access are possible. One possibility would be to provide a self-contained software, the customer can deploy in his system architecture. This can be done for example in the case of solutions for WiKi-pages where the customer installs respective software in his web server. Another possibility would be, for the OSS provider to host the solution himself and provide system architectures of customers access. This can be done for example in the case of retailers which contain systems only for managing a catalogue of available products and processing incoming orders. They could integrate an online shop solution hosted by a OSS provider through using a business connector.

@@

\section{SSI Connector}

@@

\section{OZG OSS Scenarios}

The information platform of the OZG contains information about many administration services in form of process diagrams \cite{BMI:Ergebnisse}. They serve as a resource to identify important OSS scenarios of administration processes. The following scenarios were identified:

@@

\section{OSS Connector}

\section{OZG System Architecture}
This section describes the system architecture of a member state relevant for the execution of the OZG an integration of a OSS provider. Due to a lack of detailed information, the system architecture is described by components responsible for a category of tasks. Each component is associated to data objects it processes. The definition of components and data objects are based on research about planned system architectures of member states.

@@

This chapter presents a message based integration architecture which enables access to OSS features of member states through a OSS provider. Multiple versions of an integration architecture are presented which fulfill increasingly complex requirements. Requirements are formulated as user scenarios which should be accomplishable through the OSS provider.
Each version is presented in the form of multiple flow charts, each describing sequential steps while carrying out a required scenario and a corresponding messaging infrastructure.

@@

\subsection{Requirements}
In the first version of the integration architecture, users of the OSS provider should be able to
\begin{itemize}
    \item fill in an application
    \item authenticate
    \item receive and send messages
\end{itemize}

@@

\begin{itemize}
     \item Expanding integration architecture with capabilities for OSS scenarios of different areas than online administration self-service
     \begin{itemize}
          \item online payment self-service (PayPal, psd2)
          \item online shopping self-service (Amazon)
          \item online health care self-service (DVG)
          \item online entertainment self-service (Netflix, YouTube, Spotify)
          \item online information / support self-service (StackOverflow)
     \end{itemize}
\end{itemize}
